SubsidyExplorer - Step 4: Apply proposal 'definitions' to vessel list

### 2020-04-28 scripts/05_subsidy_reform_definitions.Rmd
pre-subdirectory attempt
high seas and development status

# Path to global vessel list
vessel_list_path <- here::here("results", "3-model-parameterization", "vessel-list-2018-regions.csv")

# Paths to other data files
iuu_dat_path <- here::here("data", "raw", "fisheries", "IUU_Vessel_List", "IUUList-20190430.csv")
costello_2016_path <- here::here("data", "raw", "fisheries", "Costello_2016", "ProjectionData_2012.csv")

@@

# Directories for the output files generated by this script
results_dir <- here::here("results", "04_subsidy_reform_definitions/")
if (dir.exists(results_dir) == F) {
  dir.create(results_dir, recursive = T)
}

app_dir <- here::here("WTOtoolkit", "data/")
if (dir.exists(app_dir) == F) {
  dir.create(app_dir, recursive = T)
}

@@

```{r}
# EU countries
#Belgium, Bulgaria, Croatia, Cyprus, Denmark, Estonia, Finland, France, Germany, Greece, Ireland, Italy, Latvia, Lithuania, Malta, Netherlands, Poland, Portugal, Romania, Slovenia, Spain, Sweden, United Kingdom
eu_countries <- c("BEL", "BGR", "HRV", "CYP", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "IRL", "ITA", "LVA", "LTU", "MLT", "NLD", "POL", "PRT", "ROU", "SVN", "ESP", "SWE", "GBR")

# Dutch overseas territories: Aruba, Bonaire/St. Eustatius/Saba, Curacao, Sint Maarten
dutch_territories <- c("ABW", "BES", "CUW", "SXM")

# French overseas territories: French Guiana, French Polynesia, French Southern and Antarctic Territories, Guadeloupe, Martinique, Mayotte, New Caledonia, Reunion, Saint Barthelemy, Saint Martin, St. Pierre and Miquelon, Wallis and Futuna Islands
french_territories <- c("GUF", "PYF", "ATF", "GLP", "MTQ", "MYT", "NCL", "REU", "BLM", "MAF", "SPM", "WLF")

# Danish overseas territories: Faroe Islands, Greenland
danish_territories <- c("FRO", "GRL")

# British overseas territories: Anguilla, Ascension island, Bermuda, Cayman Islands, Falkland Islands, Guernsey, Gibraltar, British Indian Ocean Territories, Jersey, Monserratt, Pitcairn Island, South Georgia and the South Sandwich Islands, Saint Helena/Ascension/Tristan da Cunha, Tristan da Cunha, Turks and Caicos Islands, British Virgin Islands
uk_territories <- c("AIA","ASC","BMU","CYM","FLK","GGY","GIB","IOT","JEY","MSR","PCN","SGS","SHN","TAA","TCA","VGB")

eu_territories <- c(dutch_territories, french_territories, danish_territories, uk_territories)

# Norwegian overseas territories: Bouvet Islands, Svalbard and Jan Mayen Islands
norwegian_territories <- c("BVT", "SJM")

# US overseas territories:  American Samoa, Guam, Northern Marianas Islands, Puerto Rico, US Virgin Islands
us_territories <- c("ASM", "GUM", "MNP", "PRI", "VIR")

eu_northern_agreement <- c("NOR", "ISL")
```

@@

We use the flag of the vessel to assign development status. If the flag of the vessel was not able to be determined, we assume the development status to be "Developing".

```{r}
# Read in country list with development status
country_list <- read_csv(here::here("data", "country-conversion-sheet-all.csv"))

country_development_status <- country_list %>%
  group_by(iso3) %>%
  summarize(development_status_full = unique(development_status)) %>%
  mutate(development_status = str_replace(development_status_full, ' \\(.*\\)', ""),
         development_status_notes = str_extract(development_status_full, '\\(.*\\)')) %>%
  dplyr::select(iso3, development_status)

country_development_status$development_status[country_development_status$iso3 == "ATF"] <- "Developed"
country_development_status$development_status[country_development_status$iso3 == "IOT"] <- "Developed"

# Add development status to the vessel list
temp_vessel_list <- vessel_list %>%
  left_join(country_development_status, by = c("flag" = "iso3"))

# Let's also add identifies of EU vessels and US territory vessels to the vessel list
temp_vessel_list <- temp_vessel_list %>%
  mutate(is_EU = case_when(flag %in% eu_countries ~ T,
                           TRUE ~ F),
         is_EU_territory = case_when(flag %in% eu_territories ~ T,
                                     TRUE ~ F),
         is_US_territory = case_when(flag %in% us_territories ~ T,
                                     TRUE ~ F))
```

@@

In order to later be able to easily identify high seas fishing, we want to be able to easily identify the fraction of each vessel's total fishing effort spent on the high seas in a given year.

```{r}
# Get totals by vessel
vessel_totals_high_seas <- temp_vessel_list %>%
  dplyr::filter(eez_id == 0) %>%
  group_by(ssvid) %>%
  summarize(fishing_h_hs = sum(fishing_hours_eez_fao, na.rm = T),
            fishing_KWh_hs = sum(fishing_KWh_eez_fao, na.rm = T))

vessel_totals_prop_hs <- temp_vessel_list %>%
  group_by(ssvid) %>%
  summarize(fishing_h = sum(fishing_hours_eez_fao, na.rm = T),
            fishing_KWh = sum(fishing_KWh_eez_fao, na.rm = T)) %>%
  ungroup() %>%
  left_join(vessel_totals_high_seas, by = "ssvid") %>%
  mutate(prop_fishing_hours_HS = case_when(is.na(fishing_h_hs) ~ 0, 
                                           TRUE ~ fishing_h_hs/fishing_h),
         prop_fishing_KWh_HS = case_when(is.na(fishing_KWh_hs) ~ 0, 
                                         TRUE ~ fishing_KWh_hs/fishing_KWh)) %>%
  dplyr::select(ssvid, prop_fishing_hours_HS, prop_fishing_KWh_HS)

temp_vessel_list <- temp_vessel_list %>%
  left_join(vessel_totals_prop_hs, by = "ssvid")
```

>>>>

# Path to global vessel list
vessel_list_path <- here::here("results", "04-model-parameterization", "vessel_list_2018_regions.csv")

# Paths to other data files
iuu_dat_path <- here::here("data", "raw", "IUU-Vessel-List-2019", "IUUList-20190430.csv")
costello_2016_path <- here::here("data", "edited", "costello_et_al_2016_upsides_tidy.csv")

@@

# Directories for the output files generated by this script
results_dir <- here::here("results", "05-subsidy-reform-definitions/")
if (dir.exists(results_dir) == F) {
  dir.create(results_dir, recursive = T)
}

app_dir <- here::here("app", "data/")
if (dir.exists(app_dir) == F) {
  dir.create(app_dir, recursive = T)
}

@@

```{r}
# Country lookup table
country_lookup <- read_csv(here::here("data", "lookup-tables", "country_dependencies.csv"))

# Countries participating in the Northern Agreement for fishing
eu_northern_agreement <- c("NOR", "ISL")
```

@@

We use the flag of the vessel to assign development status.

```{r}
# Sovereign development statuses
sov_country_development_status <- country_lookup %>%
  dplyr::filter(!is_overseas_territory) %>%
  group_by(iso3) %>%
  summarize(development_status = case_when(!is.na(development_status) ~ development_status,
                                                TRUE ~ "Developing"))
# Get development status
ter_development_status <- country_lookup %>%
  dplyr::filter(is_overseas_territory) %>%
  dplyr::select(-development_status) %>%
  left_join(sov_country_development_status, by = c("sovereign_iso3" = "iso3")) %>%
  dplyr::select(iso3, development_status)

# combine
development_status <- sov_country_development_status %>%
  bind_rows(ter_development_status)

# Add development status to the vessel list
temp_vessel_list <- vessel_list %>%
  left_join(development_status, by = c("flag_iso3" = "iso3"))

In order to later be able to identify high seas fishing, we need to determine fraction of each vessel's total annual fishing effort spent on the high seas.

```{r}
# Get high seas totals by vessel
vessel_totals_high_seas <- temp_vessel_list %>%
  dplyr::filter(eez_id == 0) %>%
  group_by(ssvid) %>%
  summarize(fishing_h_hs = sum(fishing_hours_eez_fao_ter, na.rm = T),
            fishing_KWh_hs = sum(fishing_KWh_eez_fao_ter, na.rm = T))

# Now calculate high seas proportions
vessel_totals_prop_hs <- temp_vessel_list %>%
  group_by(ssvid) %>%
  summarize(fishing_h = sum(fishing_hours_eez_fao_ter, na.rm = T),
            fishing_KWh = sum(fishing_KWh_eez_fao_ter, na.rm = T)) %>%
  ungroup() %>%
  left_join(vessel_totals_high_seas, by = "ssvid") %>%
  mutate(prop_fishing_hours_high_seas = case_when(is.na(fishing_h_hs) ~ 0, 
                                           TRUE ~ fishing_h_hs/fishing_h),
         prop_fishing_KWh_high_seas = case_when(is.na(fishing_KWh_hs) ~ 0, 
                                         TRUE ~ fishing_KWh_hs/fishing_KWh)) %>%
  dplyr::select(ssvid, prop_fishing_hours_high_seas, prop_fishing_KWh_high_seas)

# Add back into vessel list
temp_vessel_list <- temp_vessel_list %>%
  left_join(vessel_totals_prop_hs, by = "ssvid")
```

@@
